---
title: "getcdw"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

getcdw is a tool to make it easier to query the `CADS` data warehouse from R.

You can install getcdw from github:

```{r install, eval = FALSE}
devtools::install_github("tarakc02/getcdw")
```

## Queries

You can type a query:

```{r query1, warning = FALSE, message = FALSE}
# dplyr just prints the returned data.frames prettier
library(dplyr)
library(getcdw)

get_cdw("select report_name from cdw.d_entity_mv where entity_id = 640993")
```

You can also pass a query from a file, for instance: `get_cdw("sql/filename.sql")`.

## Table and column search

Look for tables with `find_tables()`:

```{r find}
# want the committees table, but don't remember the exact name:
find_tables("committee")

# or view all of the d_bio tables:
find_tables("d_bio")

# all of the giving summary tables are like sf_something_summary_mv:
find_tables("^sf_.+_summary_mv$")
```

Or look for specific columns, either throughout the database or within a given table:

```{r find_columns}
# what's the name of the household_id column in d_entity?
find_columns("household", table_name = "d_entity_mv")

# where are all of the places degree information appears?
find_columns("degree")
```

## Code search

You can search through TMS tables to find codes:

```{r find_codes}
# where/how do we record peace_corps participation?
find_codes("peace corps")

# interest in neuroscience may present itself in terms of a major/minor, 
# an explicit interest code, attendance at an event, etc. let's look for
# all possibilities:
find_codes("neuroscience")

# i just want to look for neuro-related event codes
find_codes("neuro", "^activity$")
```
