---
title: "getcdw"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

getcdw is a tool to make it easier to query the `CADS` data warehouse from R.

You can install getcdw from github:

```{r install, eval = FALSE}
devtools::install_github("tarakc02/getcdw")
```

## Queries

You can type a query:

```{r query1, warning = FALSE, message = FALSE}
# dplyr just prints the returned data.frames prettier
library(dplyr)
library(getcdw)

get_cdw("select report_name from cdw.d_entity_mv where entity_id = 640993")
```

You can also pass a query from a file, for instance: `get_cdw("sql/filename.sql")`.

## Table and column search

Look for tables with `find_tables()`:

```{r find}
# want the committees table, but don't remember the exact name:
find_tables("committee")

# or view all of the d_bio tables:
find_tables("d_bio")

# all of the giving summary tables are like sf_something_summary_mv:
find_tables("^sf_.+_summary_mv$")
```

Or look for specific columns, either throughout the database or within a given table:

```{r find_columns}
# what's the name of the household_id column in d_entity?
find_columns("household", table_name = "d_entity_mv")

# where are all of the places degree information appears?
find_columns("degree")
```

## Code search

You can search through TMS tables to find codes:

```{r find_codes}
# where/how do we record peace_corps participation?
find_codes("peace corps")

# interest in neuroscience may present itself in terms of a major/minor, 
# an explicit interest code, attendance at an event, etc. let's look for
# all possibilities:
find_codes("neuroscience")

# i just want to look for neuro-related event codes
find_codes("neuro", "^activity$")
```

## dplyr backend

If you use dplyr, you can now query the data warehouse directly, rather than writing SQL and pulling data into a local data frame:

```{r src_oracle_demo, warning = FALSE, message = FALSE}
library(dplyr)

# each table you reference is described using the dplyr's tbl() function
entity <- tbl_cdw("cdw.d_entity_mv")
transactions <- tbl_cdw("cdw.f_transaction_detail_mv")

# note that auto-complete works inside the dplyr verbs!
entity <- entity %>% select(entity_id, person_or_org)
entity

# the top donors of 2001, along with whether they are people or organizations
transactions %>%
    filter(between(giving_record_dt,
                   to_date('01/01/2001', 'mm/dd/yyyy'),
                   to_date('12/31/2001', 'mm/dd/yyyy'))) %>%
    filter(pledged_basis_flg == "Y") %>%
    group_by(donor_entity_id_nbr) %>%
    summarise(giving = sum(benefit_dept_credited_amt)) %>%
    inner_join(entity, by = c("donor_entity_id_nbr" = "entity_id")) %>%
    arrange(desc(giving))

# dplyr will not actually pull the data into memory, which makes things
# pretty fast. when you're ready to actually download a data frame, run collect():

transactions %>%
    filter(between(giving_record_dt,
                   to_date('01/01/2001', 'mm/dd/yyyy'),
                   to_date('12/31/2001', 'mm/dd/yyyy'))) %>%
    filter(pledged_basis_flg == "Y") %>%
    group_by(donor_entity_id_nbr) %>%
    summarise(giving = sum(benefit_dept_credited_amt)) %>%
    filter(giving > 5000000) %>% 
    collect
```

## Parameterized Templates

The `parameterize_template` function allows you to turn templates that have ##highlighted## variables into functions:

```{r param_template_simple}
f <- parameterize_template("my name is ##name##")
f("tarak")

# need named arguments when more than 1 parameter:
g <-parameterize_template("hi i'm ##name##, i am ##age## years old") 
g(name = "tarak", age = 36)
```

`parameterize_template` is useful in creating reports because it allows you to write SQL with parameters, with the added benefit of the convenient function interface. 

```{r parameterized_templates}
template <- "
select donor_entity_id_nbr, 
sum(benefit_dept_credited_amt) as giving
from cdw.f_transaction_detail_mv
where pledged_basis_flg = 'Y'
and giving_record_dt between to_date('##from_date##', 'mm/dd/yyyy') 
and to_date('##to_date##', 'mm/dd/yyyy')
group by donor_entity_id_nbr
having sum(benefit_dept_credited_amt) > 5000000"

top_donors_between <- parameterize_template(template)
report <- top_donors_between(from_date = "01/01/2001", to_date = "12/31/2001")
get_cdw(report)
```

Just like with `get_cdw`, `parameterize_template` works with files also, for example `parameterize_template('sql/my-template.sql')`.
